Estimating Information Transfer in High-dimensional Linear Regression under Distribution Shifts

### 2021-07-20 revision_journal/results.tex
start intro

Our key idea is to write equation \eqref{Lbias_idea} as the derivative of a certain polynomial with respect to $x$ at $x = 0$:
\begin{align}\nonumber %\label{Lbias_idea2}
L_{\bias} (a)&\approx \|\bv_a\|^2 \cdot\left. \frac{\dd }{\dd x}\right|_{x=0}\frac1p\bigtr{  \Big( a^2(Z^{(1)})^\top Z^{(1)} + x \big((Z^{(1)})^\top Z^{(1)}\big)^2 + (Z^{(2)})^\top Z^{(2)} \Big)^{-1}  }.
\end{align}
It is well-known that the empirical spectral distributions (ESD) of $(Z^{(i)})^\top Z^{(i)}$, $i=1,2$, satisfy the famous Marchenko-Pastur (MP) law asymptotically \cite{MP}. From the MP law of $(Z^{(1)})^\top Z^{(1)}$, we can also derive the asymptotic ESD of $a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2$ for any fixed $a\in \R$ and $x>0$. Due to the rotational invariance of multivariate Gaussian distributions,
%$a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2$ and $(Z^{(2)})^\top Z^{(2)}$ are asymptotically freely independent from each other \cite{nica2006lectures}. Hence
the asymptotic ESD of $a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2+(Z^{(2)})^\top Z^{(2)}$ is given by the free additive convolution (or free addition) of the asymptotic ESD of $a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2$ and the MP law of $(Z^{(2)})^\top Z^{(2)}$\cite{nica2006lectures}.
%In particular, a sharp convergence estimate has been proved in \cite{BES_free1,BES_free2} for the free addition of two probability measures.
Finally, We use the result of \citet{BES_free1,BES_free2} to obtain a sharp estimate of
$$\frac1p\bigtr{  \Big( a^2(Z^{(1)})^\top Z^{(1)}+ x \big((Z^{(1)})^\top Z^{(1)}\big)^2+(Z^{(2)})^\top Z^{(2)} \Big)^{-1}},$$
which implies a sharp estimate of $L_{\bias}(a)$.
%Taking the derivative with respect to $x$ at $x=0$ gives the exact asymptotic limit of $L_{\bias}(a)$.
The complete proof of Theorem \ref{cor_MTL_loss} can be found in Section \ref{app_iso_cov}.

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/covariate_and_model_shift_a.eps}
		\caption{Varying sample size under covariate shift}
		\label{fig_sec3_cov_mo_a}
	\end{subfigure}
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/covariate_and_model_shift_b.eps}
		\caption{Varying sample size under model shift}
		\label{fig_sec3_cov_mo_b}
	\end{subfigure}
	\caption{We further show that our observations in Figures \ref{fig_sec31} and \ref{fig_sec3_model_shift} extend to settings with both covariate and model shift.
	Both simulation use $p = 100, n_2 = 300, \sigma = 1/2$.
	Figure \ref{fig_sec3_cov_mo_a} fixes $\mu = 0.1$ while varying $\lambda, n_1$.
	Figure \ref{fig_sec3_cov_mo_b} fixes $\lambda = 4$ while varying $\mu, n_1$.}
	\label{fig_sec33}
\end{figure}

>>>>

Our key idea is to write equation \eqref{Lbias_idea} as the derivative of a certain polynomial with respect to $x$ at $x = 0$:
\begin{align}\nonumber %\label{Lbias_idea2}
L_{\bias} (a)&\approx \|\bv_a\|^2 \cdot\left. \frac{\dd }{\dd x}\right|_{x=0}\frac1p\bigtr{  \Big( a^2(Z^{(1)})^\top Z^{(1)} + x \big((Z^{(1)})^\top Z^{(1)}\big)^2 + (Z^{(2)})^\top Z^{(2)} \Big)^{-1}  }.
\end{align}
It is well-known that the empirical spectral distributions (ESD) of $(Z^{(i)})^\top Z^{(i)}$, $i=1,2$, satisfy the famous Marchenko-Pastur (MP) law asymptotically \cite{MP}. From the MP law of $(Z^{(1)})^\top Z^{(1)}$, we can also derive the asymptotic ESD of $a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2$ for any fixed $a\in \R$ and $x>0$. Due to the rotational invariance of multivariate Gaussian distributions,
%$a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2$ and $(Z^{(2)})^\top Z^{(2)}$ are asymptotically freely independent from each other \cite{nica2006lectures}. Hence
the asymptotic ESD of $a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2+(Z^{(2)})^\top Z^{(2)}$ is given by the free additive convolution (or free addition) of the asymptotic ESD of $a^2(Z^{(1)})^\top Z^{(1)}+ x[(Z^{(1)})^\top Z^{(1)}]^2$ and the MP law of $(Z^{(2)})^\top Z^{(2)}$\cite{nica2006lectures}.
%In particular, a sharp convergence estimate has been proved in \cite{BES_free1,BES_free2} for the free addition of two probability measures.
Finally, We use the result of \citet{BES_free1,BES_free2} to obtain a sharp estimate of
$$\frac1p\bigtr{  \Big( a^2(Z^{(1)})^\top Z^{(1)}+ x \big((Z^{(1)})^\top Z^{(1)}\big)^2+(Z^{(2)})^\top Z^{(2)} \Big)^{-1}},$$
which implies a sharp estimate of $L_{\bias}(a)$.
%Taking the derivative with respect to $x$ at $x=0$ gives the exact asymptotic limit of $L_{\bias}(a)$.
\todo{can we add a few sentences to describe any technical challenge here? right now it sounds like our result is a direct corollary, which is not good?}
The complete proof of Theorem \ref{cor_MTL_loss} can be found in Section \ref{app_iso_cov}.

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/covariate_and_model_shift_a.eps}
		\caption{Fixed $\mu$, different $\lambda$}
		\label{fig_sec3_cov_mo_a}
	\end{subfigure}
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/covariate_and_model_shift_b.eps}
		\caption{Fixed $\lambda$, different $\mu$}
		\label{fig_sec3_cov_mo_b}
	\end{subfigure}
	\caption{We further show that our observations in Figures \ref{fig_sec31} and \ref{fig_sec3_model_shift} extend to settings with both covariate and model shift.
	Both simulation use $p = 100, n_2 = 300, \sigma = 1/2$.
	Figure \ref{fig_sec3_cov_mo_a} fixes $\mu = 0.1$ while varying $\lambda, n_1$.
	Figure \ref{fig_sec3_cov_mo_b} fixes $\lambda = 4$ while varying $\mu, n_1$.}
	\label{fig_sec33}
\end{figure}
